Soss: Declarative Probabilistic Programming via Runtime Code Generation

### 2020-01-06 paper/paper.tex
edits

\begin{acks}
We would like to acknowledge...
\end{acks}

@@

\begin{figure}[t]
\centering
\begin{subfigure}[t]{0.48\textwidth}
\begin{minted}{julia1}
julia> Soss.sourceRand()(m)
quote
    σ = rand(HalfNormal())
    λ = rand(HalfCauchy())
    σJ = rand(HalfNormal(λ))
    σI = rand(HalfNormal(λ))
    β = rand(iid(J, Normal(0, σJ)))
    α = rand(iid(I, Normal(0, σI)))
    yhat = α .+ β'
    y = rand(For(I, J) do i, j
                Normal(yhat[i, j], σ)
            end)
    (I=I, J=J, yhat=yhat, λ=λ, σI=σI, 
    σJ=σJ, α=α, β=β, σ=σ, y=y) 
end
\end{minted}
\caption{\footnotesize{\texttt{rand}}}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\begin{minted}{julia1}
julia> Soss.sourceLogpdf()(m)
quote
    _ℓ = 0.0
    _ℓ += logpdf(HalfNormal(), σ)
    _ℓ += logpdf(HalfCauchy(), λ)
    _ℓ += logpdf(HalfNormal(λ), σJ)
    _ℓ += logpdf(HalfNormal(λ), σI)
    _ℓ += logpdf(Normal(0, σJ) |> iid(J), β)
    _ℓ += logpdf(Normal(0, σI) |> iid(I), α)
    yhat = α .+ β'
    _ℓ += logpdf(For(I, J) do i, j
                Normal(yhat[i, j], σ)
            end, y)
    return _ℓ
end
\end{minted}
\caption{\footnotesize{\texttt{logpdf}}}
\end{subfigure}
\caption{Generated code when calling \footnotesize{\texttt{rand(m)}} and \footnotesize{\texttt{logpdf(m,x)}}.} 
\label{} 
\end{figure}

\section{Extended example }
\todo[inline]{Break this up! This is just a staging area}

\begin{minted}{julia1}
julia> using Revise, Soss, Random

julia> Random.seed!(1);

julia> m = @model I,J begin
           λ ~ HalfCauchy()
           σI ~ HalfNormal(λ)
           σJ ~ HalfNormal(λ)
           α ~ Normal(0, σI) |> iid(I)
           β ~ Normal(0, σJ) |> iid(J)
           yhat = α .+ β'
           σ ~ HalfNormal()
           y ~ For(I,J) do i,j 
                   Normal(yhat[i,j], σ)
               end
       end;

julia> truth = rand(m(I=2,J=3)); pairs(truth);

julia> post = dynamicHMC(m(I=2,J=3), (y=truth.y,));

julia> postpred = predict(m(I=2,J=3), post) |> particles;

julia> postpred.yhat
2×3 Array{Particles{Float64,1000},2}:
 0.241 ± 0.26  0.527 ± 0.3   -0.556 ± 0.38
 0.22 ± 0.26   0.506 ± 0.32  -0.577 ± 0.37

julia> postpred.y
2×3 Array{Particles{Float64,1000},2}:
 0.224 ± 0.58  0.529 ± 0.58  -0.58 ± 0.64
 0.223 ± 0.56  0.483 ± 0.58  -0.58 ± 0.61

julia> m.args
2-element Array{Symbol,1}:
 :I
 :J

julia> pairs(m.vals)
pairs(::NamedTuple) with 1 entry:
  :yhat => :(α .+ β')

julia> pairs(m.dists)
pairs(::NamedTuple) with 7 entries:
  :λ  => :(HalfCauchy())
  :σI => :(HalfNormal(λ))
  :σJ => :(HalfNormal(λ))
  :α  => :(Normal(0, σI) |> iid(I))
  :β  => :(Normal(0, σJ) |> iid(J))
  :σ  => :(HalfNormal())
  :y  => :(For(I, J) do i, j…

\end{minted}

>>>>

\begin{acks}
The authors would like to thank the Julia community, in particular Seth Axen, Tamas Papp, Fredrik Bagge Carlson, the Turing and Gen development teams, and the Julia founders and core developers. 
\end{acks}

@@

% \begin{figure}[t]
% \centering
% \begin{subfigure}[t]{0.48\textwidth}
% \begin{minted}{julia1}
% julia> Soss.sourceRand()(m)
% quote
%     σ = rand(HalfNormal())
%     λ = rand(HalfCauchy())
%     σJ = rand(HalfNormal(λ))
%     σI = rand(HalfNormal(λ))
%     β = rand(iid(J, Normal(0, σJ)))
%     α = rand(iid(I, Normal(0, σI)))
%     yhat = α .+ β'
%     y = rand(For(I, J) do i, j
%                 Normal(yhat[i, j], σ)
%             end)
%     (I=I, J=J, yhat=yhat, λ=λ, σI=σI, 
%     σJ=σJ, α=α, β=β, σ=σ, y=y) 
% end
% \end{minted}
% \caption{\footnotesize{\texttt{rand}}}
% \end{subfigure}
% \begin{subfigure}[t]{0.48\textwidth}
% \begin{minted}{julia1}
% julia> Soss.sourceLogpdf()(m)
% quote
%     _ℓ = 0.0
%     _ℓ += logpdf(HalfNormal(), σ)
%     _ℓ += logpdf(HalfCauchy(), λ)
%     _ℓ += logpdf(HalfNormal(λ), σJ)
%     _ℓ += logpdf(HalfNormal(λ), σI)
%     _ℓ += logpdf(Normal(0, σJ) |> iid(J), β)
%     _ℓ += logpdf(Normal(0, σI) |> iid(I), α)
%     yhat = α .+ β'
%     _ℓ += logpdf(For(I, J) do i, j
%                 Normal(yhat[i, j], σ)
%             end, y)
%     return _ℓ
% end
% \end{minted}
% \caption{\footnotesize{\texttt{logpdf}}}
% \end{subfigure}
% \caption{Generated code when calling \footnotesize{\texttt{rand(m)}} and \footnotesize{\texttt{logpdf(m,x)}}.} 
% \label{} 
% \end{figure}

% \section{Extended example }

% \begin{minted}{julia1}
% julia> using Revise, Soss, Random

% julia> Random.seed!(1);

% julia> m = @model I,J begin
%            λ ~ HalfCauchy()
%            σI ~ HalfNormal(λ)
%            σJ ~ HalfNormal(λ)
%            α ~ Normal(0, σI) |> iid(I)
%            β ~ Normal(0, σJ) |> iid(J)
%            yhat = α .+ β'
%            σ ~ HalfNormal()
%            y ~ For(I,J) do i,j 
%                    Normal(yhat[i,j], σ)
%                end
%        end;

% julia> truth = rand(m(I=2,J=3)); pairs(truth);

% julia> post = dynamicHMC(m(I=2,J=3), (y=truth.y,));

% julia> postpred = predict(m(I=2,J=3), post) |> particles;

% julia> postpred.yhat
% 2×3 Array{Particles{Float64,1000},2}:
%  0.241 ± 0.26  0.527 ± 0.3   -0.556 ± 0.38
%  0.22 ± 0.26   0.506 ± 0.32  -0.577 ± 0.37

% julia> postpred.y
% 2×3 Array{Particles{Float64,1000},2}:
%  0.224 ± 0.58  0.529 ± 0.58  -0.58 ± 0.64
%  0.223 ± 0.56  0.483 ± 0.58  -0.58 ± 0.61

% julia> m.args
% 2-element Array{Symbol,1}:
%  :I
%  :J

% julia> pairs(m.vals)
% pairs(::NamedTuple) with 1 entry:
%   :yhat => :(α .+ β')

% julia> pairs(m.dists)
% pairs(::NamedTuple) with 7 entries:
%   :λ  => :(HalfCauchy())
%   :σI => :(HalfNormal(λ))
%   :σJ => :(HalfNormal(λ))
%   :α  => :(Normal(0, σI) |> iid(I))
%   :β  => :(Normal(0, σJ) |> iid(J))
%   :σ  => :(HalfNormal())
%   :y  => :(For(I, J) do i, j…

% \end{minted}
